---
title: "Initial Steps for Preprocessing"
output: html_notebook
---

## Introduction

This document contains all the required steps for loading, preprocessing, and preparing the data for analysis. Functions and helper scripts are included here for standalone execution. It includes: - Dynamic cancer and layer selection. - Covariate filtering and encoding. - Feature filtering, imputation, and layer combination. - Splitting data into train and test sets.

------------------------------------------------------------------------

## Required Libraries

```{r setup, message=FALSE, warning=FALSE}
# Load required libraries
library(SummarizedExperiment)
library(caret)
library(ggplot2)
library(impute)
```

------------------------------------------------------------------------

## Helper Functions

```{r}
# Get a list of RangedSummarizedExperiment and SummarizedExperiment objects
getExperimentsList <- function(paths, featuresToEnsure, removeHPVPositive = FALSE) {
  experimentList <- list()
  for (i in 1:length(paths)) {
    layerName <- names(paths)[i]

    # Check if the file exists before reading
    if (!file.exists(paths[[layerName]])) {
      stop(paste("File does not exist for layer:", layerName))
    }

    experiment <- readRDS(paths[[layerName]])

    # Ensure required features exist
    missingFeatures <- setdiff(featuresToEnsure, colnames(SummarizedExperiment::colData(experiment)))
    if (length(missingFeatures) > 0) {
      stop(paste("Missing required features in experiment:", layerName, "-", paste(missingFeatures, collapse = ", ")))
    }

    # Remove samples with missing metadata
    for (feature in featuresToEnsure) {
      experiment <- experiment[, !is.na(experiment[[feature]])]
    }

    # Remove HPV-positive samples if requested
    if (removeHPVPositive) {
      if ("HPV.status" %in% colnames(SummarizedExperiment::colData(experiment))) {
        experiment <- experiment[, experiment$HPV.status == "Negative"]
      } else {
        warning(paste("HPV.status column is missing for layer:", layerName))
      }
    }

    # Convert Age to numeric if it exists
    if ("Age" %in% colnames(SummarizedExperiment::colData(experiment))) {
      experiment$Age <- as.numeric(experiment$Age)
    } else {
      warning(paste("Age column is missing for layer:", layerName))
    }

    experimentList[[layerName]] <- experiment
  }
  return(experimentList)
}

# Get paths dynamically for each cancer and layer
getPathsByLayerAndCancer <- function(getLayers = TRUE) {
  layerNames <- c("miRNA", "RNAseq_filtered", "methylation", "RPPA", "RNAseq_NORMAL_filtered", "binary-mutation", "SCNA")
  layerPartialPaths <- c("miR", "RNAseq", "methylation", "RPPA", "normal/RNAseq", "mutation", "scna")
  commonPath <- "/restricted/projectnb/agedisease/CBMrepositoryData/TCGA-GDC/"
  filteredPath <- "/processed_data/filtered/TCGA-*_"
  allPath <- "/processed_data/all/TCGA-*"
  pathsByLayerAndCancer <- list()

  for (i in 1:length(layerNames)) {
    # Construct path for the layer
    filePath <- paste(commonPath, layerPartialPaths[i], sep = "")
    filePath <- if (i < 6) paste(filePath, filteredPath, sep = "") else paste(filePath, allPath, sep = "")
    filePath <- paste(filePath, layerNames[i], ".rds", sep = "")

    # Get paths and log if no files are found
    layerPaths <- Sys.glob(filePath)
    if (length(layerPaths) == 0) {
      warning(paste("No files found for layer:", layerNames[i]))
    }
    pathsByLayerAndCancer[[layerNames[i]]] <- layerPaths
  }
  return(pathsByLayerAndCancer)
}

# Extract cancer names from paths
extractCancerNames <- function(pathVector) {
  # Extract cancer type (e.g., "TCGA-HNSC") from the filename
  return(sapply(pathVector, function(x) {
    match <- regexpr("TCGA-[A-Z]+_", basename(x))  # Match patterns like "TCGA-HNSC_"
    if (match > 0) {
      return(substr(basename(x), match, match + attr(match, "match.length") - 2))  # Remove the trailing "_"
    } else {
      return(NA)
    }
  }))
}

# Impute missing values using KNN
imputeMissingValues <- function(experimentsList, assayNames, needsImputation) {
  for (i in 1:length(experimentsList)) {
    layerName <- names(experimentsList)[i]
    if (needsImputation[[layerName]]) {
      experiment <- experimentsList[[layerName]]
      assayName <- assayNames[[layerName]]

      # Check if assay exists
      if (!assayName %in% names(SummarizedExperiment::assays(experiment))) {
        warning(paste("Assay not found for imputation in layer:", layerName))
        next
      }

      # Impute only if there are missing values
      if (any(is.na(SummarizedExperiment::assay(experiment, assayName)))) {
        data_imputed <- impute::impute.knn(as.matrix(SummarizedExperiment::assay(experiment, assayName)), k = 5)$data
        SummarizedExperiment::assay(experiment, paste(assayName, "-imputed", sep = "")) <- data_imputed
        experimentsList[[layerName]] <- experiment
      } else {
        message(paste("No missing values found for layer:", layerName))
      }
    }
  }
  return(experimentsList)
}

# Filter for covariates
filterForCovariates <- function(experimentsList, columns_to_keep) {
  covariatesList <- list()
  for (i in 1:length(experimentsList)) {
    experiment <- experimentsList[[i]]

    # Check if columns exist
    missingColumns <- setdiff(columns_to_keep, colnames(SummarizedExperiment::colData(experiment)))
    if (length(missingColumns) > 0) {
      stop(paste("Missing required columns in experiment:", paste(missingColumns, collapse = ", ")))
    }

    covariates <- SummarizedExperiment::colData(experiment)[, columns_to_keep, drop = FALSE]
    covariates <- covariates[complete.cases(covariates), ]
    experimentsList[[i]] <- experiment[, rownames(covariates)]
    covariatesList <- c(covariatesList, covariates)
  }
  return(list(experiments = experimentsList, covariates = covariatesList))
}

# Get cleaned assay data
getCleanedAssayData <- function(experiment, assayName, simplifiedDataPairs = list(), useImputed = FALSE) {
  if (length(simplifiedDataPairs) > 0) {
    simplifiedData <- simplifiedDataPairs[[1]]$data
    significantFeatures <- simplifiedData[simplifiedData[[simplifiedDataPairs[[1]]$metric]] <= 0.10, ]
    if (nrow(significantFeatures) == 0) {
      warning("No significant features found.")
      return(NULL)
    }
    if (assayName == "M-values") {
      experiment <- experiment[rownames(significantFeatures), ]
    } else {
      experiment <- experiment[rowRanges(experiment)$gene_name %in% rownames(significantFeatures), ]
    }
  }
  assayType <- if (useImputed) paste(assayName, "-imputed", sep = "") else assayName
  if (!assayType %in% names(SummarizedExperiment::assays(experiment))) {
    stop(paste("Assay type not found:", assayType))
  }
  assayData <- as.data.frame(SummarizedExperiment::assay(experiment, assayType))
  return(t(assayData))
}

# One-hot encode covariates
encode_covariates <- function(covariates, assayData, encodingHPV = FALSE) {
  if (!identical(rownames(covariates), rownames(assayData))) {
    stop("Samples of covariates and assay data do not match!")
  }
  
  # Dynamically build the formula based on available columns
  availableColumns <- colnames(covariates)
  baseFormula <- "~ Age"
  if ("gender" %in% availableColumns) baseFormula <- paste(baseFormula, "+ gender")
  if ("race" %in% availableColumns) baseFormula <- paste(baseFormula, "+ race")
  if (encodingHPV && "HPV.status" %in% availableColumns) baseFormula <- paste(baseFormula, "+ HPV.status")
  
  formula <- as.formula(baseFormula)
  
  # Check if at least one categorical variable exists
  categoricalVars <- setdiff(colnames(model.matrix(formula, covariates)), "(Intercept)")
  if (length(categoricalVars) == 0) {
    stop("No categorical variables found for encoding!")
  }
  
  dummies <- caret::dummyVars(formula, data = covariates)
  encoded_covariates <- data.frame(predict(dummies, newdata = covariates))
  encoded_covariates$Age <- covariates$Age
  return(cbind(encoded_covariates, assayData))
}


# Filter features
filterFeatures <- function(assayData, experiment, relativeVariance = TRUE, highCorrelationToEachOther = TRUE, cutoffThreshold = 0.01) {
  featureVariance <- apply(assayData, 1, var, na.rm = TRUE)
  selectedFeatures <- names(featureVariance[featureVariance > cutoffThreshold])
  filteredData <- assayData[selectedFeatures, , drop = FALSE]
  if (highCorrelationToEachOther && ncol(filteredData) > 1) {
    correlationMatrix <- cor(t(filteredData), use = "pairwise.complete.obs")
    highCorrelationPairs <- which(abs(correlationMatrix) > 0.9, arr.ind = TRUE)
    uniqueFeatures <- unique(highCorrelationPairs[, 1])
    filteredData <- filteredData[-uniqueFeatures, , drop = FALSE]
  }
  return(filteredData)
}


splitTrainTest <- function(data, trainSetSize = 0.8, seed = 42) {
  set.seed(seed)
  
  # Ensure rownames are unique to avoid conflicts
  if (any(duplicated(rownames(data)))) {
    stop("Row names of the dataset are not unique, which may cause issues during splitting.")
  }
  
  # Create an index for training samples
  trainIndex <- sample(1:nrow(data), size = round(nrow(data) * trainSetSize), replace = FALSE)
  
  # Split the data
  trainData <- data[trainIndex, , drop = FALSE]
  testData <- data[-trainIndex, , drop = FALSE]
  
  return(list(train = trainData, test = testData))
}

```

------------------------------------------------------------------------

## Dynamic Preprocessing Function

```{r}
dynamicPreprocess <- function(
  cancerName,
  layers,
  covariatesToKeep,
  featureFilterOptions = list(varianceThreshold = 0.01, topFeatures = 100),
  imputeData = TRUE,
  combineLayers = FALSE,
  encodeCovariates = TRUE,
  splitSize = 0.8,
  seed = 42
) {
  # Get all paths by layer and cancer dynamically
  pathsByLayer <- getPathsByLayerAndCancer(getLayers = TRUE)
  
  # Initialize storage for paths, assay names, and experiments
  layerPaths <- list()
  assayNames <- list()
  experimentsList <- list()
  needsImputation <- list()

  # Extract valid cancer names dynamically from the paths of the first requested layer
  validCancers <- extractCancerNames(pathsByLayer[[layers[1]]])
  cancerIndex <- match(cancerName, validCancers)
  if (is.na(cancerIndex)) {
    stop(paste("Invalid cancer name:", cancerName, ". Valid options are:", paste(validCancers, collapse=", ")))
  }

  # Dynamically process layers
  for (layer in layers) {
    layerPaths[[layer]] <- pathsByLayer[[layer]]
    if (is.null(layerPaths[[layer]]) || length(layerPaths[[layer]]) == 0) {
      stop(paste("No paths available for layer:", layer))
    }
    assayNames[[layer]] <- ifelse(layer == "methylation", "M-values", "DESeq2_log")
    needsImputation[[layer]] <- (layer == "methylation")
  }

  # Load experiments dynamically for each layer
  for (layer in names(layerPaths)) {
    layerPath <- layerPaths[[layer]][cancerIndex]  # Match the correct path for the cancer
    experiments <- getExperimentsList(
      paths = list(layer = layerPath),
      featuresToEnsure = covariatesToKeep
    )
    
    # Ensure experiments are SummarizedExperiment objects
    if (!all(sapply(experiments, inherits, "SummarizedExperiment"))) {
      stop(paste("Invalid data format for layer:", layer, "- Expected SummarizedExperiment objects."))
    }
    experimentsList[[layer]] <- experiments[[1]]  # Extract the actual experiment
  }

  # Impute missing data if enabled
  if (imputeData) {
    experimentsList <- imputeMissingValues(experimentsList, assayNames, needsImputation)
  }

  # Filter for selected covariates
  filteredDataList <- list()
  for (layer in names(experimentsList)) {
    filteredDataList[[layer]] <- filterForCovariates(
      list(experimentsList[[layer]]), # Wrapping in a list to match function signature
      columns_to_keep = covariatesToKeep
    )
  }

  # Clean and filter assay data
  cleanedAssayDataList <- list()
  for (layer in names(filteredDataList)) {
    cleanedAssayDataList[[layer]] <- getCleanedAssayData(
      experiment = filteredDataList[[layer]]$experiments[[1]],
      assayName = assayNames[[layer]],
      useImputed = imputeData  # Pass the imputeData flag to decide the assay type
    )
  }


  # Feature filtering based on variance or other criteria
  filteredFeaturesList <- list()
  for (layer in names(cleanedAssayDataList)) {
    filteredFeaturesList[[layer]] <- filterFeatures(
      assayData = cleanedAssayDataList[[layer]],
      experiment = experimentsList[[layer]],
      relativeVariance = featureFilterOptions$varianceThreshold > 0,
      highCorrelationToEachOther = FALSE,
      cutoffThreshold = featureFilterOptions$varianceThreshold
    )
  }

  # Combine layers if requested
  if (combineLayers) {
    combinedData <- combineLayers(filteredFeaturesList)
  }

  # Encode covariates if requested
  encodedDataList <- list()
  for (layer in names(filteredDataList)) {
    encodedDataList[[layer]] <- encode_covariates(
      covariates = filteredDataList[[layer]]$covariates[[1]],
      assayData = filteredFeaturesList[[layer]],
      encodingHPV = encodeCovariates
    )
  }

  # Split data into training and testing sets
  trainTestSplitList <- list()
  for (layer in names(encodedDataList)) {
    trainTestSplitList[[layer]] <- splitTrainTest(
      encodedDataList[[layer]],
      trainSetSize = splitSize,
      seed = seed
    )
  }

  # Return preprocessed data
  return(list(
    train = lapply(trainTestSplitList, function(x) x$train),
    test = lapply(trainTestSplitList, function(x) x$test),
    combinedData = if (combineLayers) combinedData else NULL
  ))
}


```

------------------------------------------------------------------------

## Example Usage

```{r}
cancerName <- "TCGA-HNSC"
layers <- c("RNAseq_filtered", "methylation")
covariatesToKeep <- c("Age", "gender", "HPV.status")
featureFilterOptions <- list(varianceThreshold = 0.01, topFeatures = 50)

# Run preprocessing
preprocessedData <- dynamicPreprocess(
  cancerName = cancerName,
  layers = layers,
  covariatesToKeep = covariatesToKeep,
  featureFilterOptions = featureFilterOptions,
  imputeData = FALSE,
  combineLayers = FALSE,
  encodeCovariates = TRUE
)

# Output results
print(preprocessedData$train)
print(preprocessedData$test)

```

------------------------------------------------------------------------
