---
title: "R Notebook"
output: html_notebook
---

```{r message=FALSE, warning=FALSE, include=FALSE}
library(e1071) 
library(randomForest) 
library(SummarizedExperiment)
library(Biobase)
library(ggplot2)
library(caret)
#library(SEtools)
library ("dplyr")
if (!file.exists("preprocessing.R")) {
  knitr::purl("preprocessing.Rmd", output = "preprocessing.R")
}
source("preprocessing.R")
source("PanClockHelperFunctions.R")
source("run_analysis_pipeline.R")
```

```{r}
# Confirm required functions exist in the environment
ls()  # Lists all objects in the environment
```

### Function to apply systematic bias correction

```{r}
bias_correction <- function(predicted_age, true_age) {
  print("Applying bias correction...")
  
  # Create a data frame for calculations
  DF <- data.frame(predicted_age = predicted_age, true_age = true_age)
  
  # Fit a linear model: predicted_age ~ true_age
  age_fit <- lm(predicted_age ~ true_age, data = DF)
  
  # Apply the bias correction formula
  DF <- DF |>
    dplyr::rename(chronological = true_age) |>
    dplyr::mutate(delta_age_corrected = predicted_age - age_fit$fitted.values) |>
    dplyr::mutate(predicted_corrected = delta_age_corrected + chronological)
  
  return(DF$predicted_corrected)
}
```

### Split data into training and testing sets

```{r}
splitTrainTest <- function(assayData, trainSetSize=0.8, seed=42){
  print("Train test splitting...")
  set.seed(seed)
  # assayData$Age <- yeo.johnson(assayData$Age, lambda=1.5)
  train_index <- caret::createDataPartition(assayData$Age, p = trainSetSize, list = FALSE)
  meta_trn <- assayData[train_index, ] 
  meta_tst <- assayData[-train_index, ]
  return(list(train = meta_trn, test = meta_tst))
}
```

### Build ML model; default is ElasticNet

```{r}
train_test_wrapper <- function(meta_trn, meta_tst, methodName="glmnet", do_plot=TRUE, seed=42, graphTitle="", stratifying=FALSE, applyBiasCorrection=TRUE) {
  print("Training model...")
  print(methodName)
  
  # Model selection using 10-fold CV
  if (!is.null(seed)) { set.seed(seed) }
  cv10 <- trainControl(method = "cv", number = 10)
  complexControl <- trainControl(method="boot")
  cv3 <- trainControl(method = "cv", number = 3)
  meta_trn <- subset(meta_trn, select=-submitter_id)
  
  # Random Forest
  if (methodName == "ranger"){
    meta_model <- caret::train(
      Age ~ .,
      data = meta_trn,
      method = methodName,
      trControl = cv10,
    )
  }
  
  # Neural Network
  else if (methodName == "nnet"){
    meta_trn$Age_bin <- cut(meta_trn$Age, breaks=5, labels=FALSE) # 5 bins
    # cv3 <- trainControl(method = "cv", number = 3, sampling = "up", index = createResample(meta_trn$Age_bin, times = 5))
    meta_trn$Age_bin <- as.factor(meta_trn$Age_bin)
    upSampled <- caret::upSample(x=as.data.frame(meta_trn[, -ncol(meta_trn)]), y=meta_trn$Age_bin)
    upSampled$Age_bin <- NULL
    # weights <- 1 / table(cut(meta_trn$Age, breaks = 5))[cut(meta_trn$Age, breaks=5)]
    meta_model <- caret::train(
      Age ~ .,
      # data = meta_trn,
      data = upSampled,
      method = methodName,
      trControl = cv3,
      # tuneLength = 5,
      linout = TRUE,
      MaxNWts = 10000,
      # trace = FALSE,
      maxit = 100,
      tuneGrid = expand.grid(
    size = c(5, 10, 20),    # Number of neurons in the hidden layer
    decay = c(0.001, 0.01, 0.1) # Regularization parameters
      ),
      preProcess = c('scale','pca'),
      metric = "RMSE",
      # weights = weights
    )
  }
  
  # Anything else
  else{
    if (stratifying){
      print("Stratifying...")
      meta_trn$Age_bin <- cut(meta_trn$Age, breaks=5, labels=FALSE)
      # cv3 <- trainControl(method = "cv", number = 3, sampling = "up", index = createResample(meta_trn$Age_bin, times = 5))
      meta_trn$Age_bin <- as.factor(meta_trn$Age_bin)
      sampled <- upSample(x=meta_trn[, -ncol(meta_trn)], y=meta_trn$Age_bin)
      sampled$Age_bin <- NULL
      sampled$Class <- NULL
      trainData <- sampled
    }
    else{
      trainData <- meta_trn
    }
    
    meta_model <- caret::train(
      Age ~ .,
      data = trainData,
      method = methodName,
      trControl = cv10,
      # preProcess = c('scale','pca'),
      # preProcess = c('center', 'scale', 'YeoJohnson', 'nzv'),
      # metric = "RMSE",
      tuneLength = 5
    )
  }
  
  print("Predicting...")

  # Prediction
  meta_age_estimate <- predict(meta_model, subset(meta_tst, select=-submitter_id))

  # Augment metadata with predicted age
  meta_tst$predicted_age <- meta_age_estimate
  R2 <- cor(meta_tst$Age, meta_tst$predicted_age)^2
  RMSE <- sqrt(mean((meta_tst$Age - meta_tst$predicted_age)^2))
  meta_tst$R2 <- R2
  meta_tst$RMSE <- RMSE

  # Plot results using graphAgeModel
  if (do_plot && nchar(graphTitle) > 0) {
    graphAgeModel(meta_tst, graphTitle, paste("R^2:", toString(R2), "- RMSE:", toString(RMSE)))
  }

  # Apply bias correction if enabled
  if (applyBiasCorrection) {
    corrected_predicted_age <- bias_correction(meta_tst$predicted_age, meta_tst$Age)
    meta_tst$predicted_corrected_age <- corrected_predicted_age
    R2_after <- cor(meta_tst$Age, meta_tst$predicted_corrected_age)^2
    RMSE_after <- sqrt(mean((meta_tst$Age - meta_tst$predicted_corrected_age)^2))
    
    if (do_plot && nchar(graphTitle) > 0) {
      graphAgeModel(
        meta_tst,
        paste("After Bias Correction: ", graphTitle, sep = ""),
        paste("R^2:", toString(round(R2_after, 3)), "- RMSE:", toString(round(RMSE_after, 3))),
        biasCorrection = TRUE)
    }
  }
  

  # Return model and augmented data
  return(list(model = meta_model, predicted = meta_tst))
}
```

### Primary wrapper function to be run by other files - make sure to impute beforehand!

# Returns: list object with indices "model" and "predictions"

# Access coefficients through listName[["model"]]$finalModel, true age with listName[["predicted"]]$age, predicted age with listName[["predicted"]]\$predicted_age

---
<!-- params: -->
<!--   inputList$experimentsList: list object of SummarizedExperiments-->
<!--   inputList$assayNames: list object of the preferred assay names -->
<!--   inputList$needsImputation: list object of booleans determining whether or not imputation is performed for each layer -->
<!--   inputList$simplifiedData: list object with "data" as index to DataFrame of file containing HNSC summary data and "metric" as index to column name corresponding to summary statistic of feature significance -->
<!--   cancerName: the abbreviation for the tumor type in question -->
<!--   splitSize: proportion of subjects to go into training set -->
<!--   testOnCompleteData: whether to use the training set also as the test set -->
<!--   methodNames: vector of ML methods to use. If more than one will perform mixture model -->
<!--   usingImputed: set to TRUE if creating new assay based on previously imputed values of old assay -->
<!--   seed: set to any integer to control randomness -->
<!--   pca: set to TRUE to perform PCA. Not currently used except for nnet -->
<!--   columnsToKeep: vector of metadata column names to keep in experiments as covariates or predictors -->
<!--   stratifying: set to TRUE to stratify subjects by age -->
---

```{r}
ModelBuilding <- function(inputList, cancerName="HNSC", splitSize=0.8, testOnCompleteData=FALSE, methodNames=c("glmnet"), graphTitle="", seed=42, pca=FALSE, columnsToKeep=c("Age", "gender", "race", "HPV.status"), stratifying=FALSE, iterationCount=NULL, applyBiasCorrection=TRUE){

  experimentsList <- inputList$experimentsList
  assayNamesList <- inputList$assayNames
  needsImputation <- inputList$needsImputation
  simplifiedDataPairs <- inputList$simplifiedData

  if (splitSize > 0.98){ testOnCompleteData <- TRUE }
  for (name in names(experimentsList)){
    if (name %in% c("CESC", "OV", "PRAD", "TGCT", "UCEC", "UCS")){ columnsToKeep <- columnsToKeep[! columnsToKeep %in% "gender"] }
  }
  if (any(names(experimentsList) != "HNSC")){ columnsToKeep <- columnsToKeep[! columnsToKeep %in% "HPV.status"]}
  if (cancerName %in% c("CESC", "OV", "PRAD", "TGCT", "UCEC", "UCS")){ columnsToKeep <- columnsToKeep[! columnsToKeep %in% "gender"] }
  if (cancerName != "HNSC"){ columnsToKeep <- columnsToKeep[! columnsToKeep %in% "HPV.status"]}
  encodingHPV <- "HPV.status" %in% columnsToKeep
  encodingGender <- "gender" %in% columnsToKeep
  
  ## Pre-processing
  experimentsAndCovariatesList <- filterForCovariates(experimentsList, columns_to_keep=columnsToKeep)
  assayList <- getCleanedAssayData(experimentsAndCovariatesList[["experiments"]], assayNamesList, needsImputation, simplifiedDataPairs=simplifiedDataPairs)
  transformedData <- filterFeatures(assayList, experimentsAndCovariatesList[["experiments"]], highCorrelationToEachOther=FALSE, relativeVariance=FALSE)
  transformedData <- combineAssays(transformedData)
  assayDataWithCovariates <- encode_covariates(experimentsAndCovariatesList[["covariates"]][[1]], transformedData, encodingHPV=encodingHPV, encodingGender=encodingGender)

  ## Running model
  train_testList <- splitTrainTest(assayDataWithCovariates, trainSetSize=splitSize, seed=seed)

    # preProcess <- preProcess(train_testList[["train"]], method = "pca", pcaComp = 100) # Reduce to 100 components
  if (testOnCompleteData){ testSet <- train_testList[["train"]] }
  else { testSet <- train_testList[["test"]] }
  
  if (length(methodNames) == 1 && (is.null(iterationCount) || iterationCount < 2)){
    print("Testing only one model!")
    # Pass `applyBiasCorrection` directly to `train_test_wrapper`
    return(train_test_wrapper(train_testList[["train"]], testSet, seed=seed, methodName=methodNames[1], graphTitle=graphTitle, stratifying=stratifying, applyBiasCorrection = applyBiasCorrection))
  }
  
  ## Mixture model
  if (length(methodNames) > 1){

    modelsList <- list()
    weightsSurvival <- c()
  
    # Creating model for each inputted method name
    startingWeight <- 1 / length(methodNames)
    for (j in 1:length(methodNames)){
      weightsSurvival <- c(weightsSurvival, startingWeight)
      modelsList[[methodNames[j]]] <- list()
    }
  
    # For each iteration get a new train test split of the 80% available
    for (j in 1:iterationCount){
      seed <- seed + 1
      train_testInitial <- splitTrainTest(train_testList$train, trainSetSize=0.8, seed=seed)
      # For each ML model
      for (k in 1:length(methodNames)){
        correlations <- c()
        methodName <- methodNames[k]
        modelPredictionPair <- train_test_wrapper(train_testInitial$train, train_testInitial$test, seed=seed, methodName=methodName, do_plot=FALSE, stratifying=stratifying, applyBiasCorrection = FALSE)
        modelsList[[methodName]][[j]] <- modelPredictionPair$model
        predictions <- modelPredictionPair$predicted
        df <- data.frame(predicted_age=predictions$predicted_age, submitter_id=predictions$submitter_id)
        # stats <- run_analysis_pipeline(experimentsList[[1]], df)
        # Weights = priors * new weights
        # weightsSurvival[k] <- weightsSurvival[k] * stats$interaction_summary$logtest[1]
        # weightsSurvival[k] <- weightsSurvival[k] * stats$interaction_summary$concordance[1]
        correlations <- cor(predictions$Age, predictions$predicted_age)
        weightsSurvival[k] <- weightsSurvival[k] * correlations^2
      }
      
      # Establishing weights for each model
      # weights <- correlations^2
      # weights <- weights / sum(weights) # Normalize to add to 1
      weightsSurvival <- weightsSurvival / sum(weightsSurvival) # Normalize to add to 1
      print(weightsSurvival)
      remainingTests <- train_testList$test
    }
  
    # Combining models based on weights to output one set of predictions
    print("Combining ensemble predictions...")
    combinedPredictions <- numeric(length(remainingTests$Age))
    modelWeightString <- ""
    for (i in 1:length(modelsList)){
      model <- modelsList[[i]]
      modelWeight <- weightsSurvival[[i]]
      currentModelPredictions <- numeric(length(remainingTests$Age))
      for (j in 1:iterationCount){
        currentModelPredictions <- currentModelPredictions + modelWeight * predict(model[[j]], subset(remainingTests, select=-submitter_id))
      }
      combinedPredictions <- combinedPredictions + currentModelPredictions
      modelWeightString <- paste(modelWeightString, ", ", names(modelsList)[i], "- Weight: ", toString(round(modelWeight, digits=3)), sep="")
    }

    R2 <- round(cor(remainingTests$Age, combinedPredictions)^2, digits=3)
    RMSE <- round(sqrt(mean((remainingTests$Age - combinedPredictions)^2)), digits=3)
    remainingTests$predicted_age <- combinedPredictions
    
    # Plot results
    caption <- paste("R^2: ",toString(R2),", RMSE: ",toString(RMSE), modelWeightString, sep="")
    graphAgeModel(remainingTests, graphTitle, caption)
  }
  
  ## Cross validation
  else if (iterationCount > 1){
    print("Cross validating...")
    seed <- seed^3 # So that if we increment the seed by 1 to test we aren't overlapping with the previous seed
    modelsList <- list()
    
    # For each iteration get a new train test split of the 80% available
    for (i in 1:iterationCount){
      seed <- seed + 1
      train_testInitial <- splitTrainTest(train_testList$train, trainSetSize=0.8, seed=seed)
      modelsList[[i]] <- train_test_wrapper(train_testInitial$train, train_testInitial$test, seed=seed, methodName=methodNames[1], do_plot=FALSE, stratifying=stratifying)
      modelsList[[i]]$R2 <- round(cor(train_testInitial$test$Age, modelsList[[i]]$predicted$predicted_age)^2, digits=3)
      modelsList[[i]]$RMSE <- round(sqrt(mean((train_testInitial$test$Age - modelsList[[i]]$predicted$predicted_age)^2)), digits=3)
    }
    
    # Make prediction on remaining data based off all the models
    remainingTests <- train_testList$test
    combinedPredictions <- numeric(length(remainingTests$Age))
    weight <- 1 / iterationCount
    for (i in 1:iterationCount){
      combinedPredictions <- combinedPredictions + weight * predict(modelsList[[i]]$model, subset(remainingTests, select=-submitter_id))
    }
    R2 <- round(cor(remainingTests$Age, combinedPredictions)^2, digits=3)
    RMSE <- round(sqrt(mean((remainingTests$Age - combinedPredictions)^2)), digits=3)
    
    remainingTests$predicted_age <- combinedPredictions
    caption <- paste("R^2: ", toString(R2), ", RMSE: ", toString(RMSE), sep="")
    graphAgeModel(remainingTests, title=graphTitle, caption=caption)
    return(list(model=modelsList, predicted=remainingTests, experimentsList=experimentsList))
  }
  else{ return(NULL) }

  # Apply bias correction to ensemble predictions if enabled
  if (applyBiasCorrection) {
    correctedPredictions <- bias_correction(combinedPredictions, remainingTests$Age)
    remainingTests$predicted_corrected_age <- correctedPredictions

    # Plot corrected results
    R2_corrected <- cor(remainingTests$Age, correctedPredictions)^2
    RMSE_corrected <- sqrt(mean((remainingTests$Age - correctedPredictions)^2))

    title <- paste("Corrected Ensemble Predictions -", graphTitle)
    caption <- paste("R^2: ", toString(round(R2_corrected, 3)), "- RMSE: ", toString(round(RMSE_corrected, 3)))
    graphAgeModel(remainingTests, title, caption, biasCorrection=TRUE)
  }
  
  return(list(model=modelsList, predicted=remainingTests, weights=weightsSurvival, experimentsList=experimentsList))
}
```

# Function to get lists containing paths for each omics layer, names for the best assay type respective to each layer, and whether or not imputation is required for each layer

```{r}
setup1 <- function(layersVector, cancerName="HNSC"){
  paths <- getPathsByLayerAndCancer()
  layerPaths <- list()
  cancerPaths <- list()
  assayNames <- list()
  needsImputation <- list()
  cancerSpecificLayerPaths <- getLayerPathsForSpecificCancer(paths, cancerName)
  
  if ("methylation" %in% layersVector) { 
    layerPaths$methylation <- cancerSpecificLayerPaths$methylation
    assayNames$methylation <- "M-values"
    needsImputation$methylation <- TRUE
  }
  if ("RNAseq" %in% layersVector) { 
    layerPaths$RNAseq <- cancerSpecificLayerPaths$RNAseq_filtered
    assayNames$RNAseq <- "DESeq2_log"
    needsImputation$RNAseq <- TRUE
  }
  if ("miRNA" %in% layersVector) {
    layerPaths$miRNA <- cancerSpecificLayerPaths$miRNA
    assayNames$miRNA <- "DESeq2_log"
    needsImputation$miRNA <- FALSE
  }
  if ("RPPA" %in% layersVector) { 
    layerPaths$RPPA <- cancerSpecificLayerPaths$RPPA
    assayNames$RPPA <- "counts"
    needsImputation$RPPA <- TRUE
  }
  
  return(list(layerPaths=layerPaths, assayNames=assayNames, needsImputation=needsImputation, cancerName=cancerName))
}
```

# Perform second part of setup, including imputing, getting significant features

```{r}
setup2 <- function(setup1List, featuresToEnsure=c("Age"), existingExperimentsList=NULL){
  PATH <- file.path( Sys.getenv("PCANAGE","/restricted/projectnb/agedisease/projects/pancancer_aging_pbock")) 
  layerPaths <- setup1List$layerPaths
  assayNames <- setup1List$assayNames
  needsImputation <- setup1List$needsImputation
  cancerName <- setup1List$cancerName
  removeHPVPositive <- FALSE
  # Set up reduced feature set based on association with age
  simplifiedData <- list()
  if (cancerName == "HNSC"){
    featuresToEnsure <- c("Age", "HPV.status")
    removeHPVPositive <- TRUE
  }

  # Get significant features for RNAseq and methylation
  if (!is.null(layerPaths$methylation)) { 
    simplifiedData$methylation <- list(data=readRDS(paste(PATH,"/results/methylation/DMP_analysis/TCGA-",cancerName,"_CpG-Sites_hg19_limma.rds", sep="")), metric="adj.P.Val")
  }
  if (!is.null(layerPaths$RNAseq)) { 
    simplifiedData$RNAseq <- list(data=readRDS(paste(PATH,"/results/RNAseq/DGE_analysis/TCGA-",cancerName,"_filtered_DESeq2_results.rds", sep="")), metric="padj")
  }
  if (!is.null(layerPaths$miRNA)) { simplifiedData$miRNA <- NULL }
  if (!is.null(layerPaths$RPPA)) { simplifiedData$RPPA <- NULL }
  
  # Get list of SummarizedExperiments for each omics layer of the cancer
  if (is.null(existingExperimentsList)){ 
    experimentsList <- getExperimentsList(layerPaths, featuresToEnsure=featuresToEnsure, removeHPVPositive=removeHPVPositive)
  }
  
  # Perform imputation
  if (is.null(existingExperimentsList)) { experimentsList <- imputeMissingValues(experimentsList, assayNames, needsImputation) } # Run if missing values in assay
  else { experimentsList <- existingExperimentsList }
  return(list(experimentsList=experimentsList, assayNames=assayNames, needsImputation=needsImputation, simplifiedData=simplifiedData))
}
```

# Wrapper for the two setup functions

```{r}
performSetup <- function(layersVector, cancerName="HNSC", featuresToEnsure=c("Age"), removeHPVPositive=FALSE, existingExperimentsList=NULL){
  return(setup2(setup1(layersVector, cancerName=cancerName), featuresToEnsure=featuresToEnsure, existingExperimentsList=existingExperimentsList))
}
```

# Wrapper for ModelBuilding with default params to make life easier. Just need to pass in vector of omics layer names and the cancer name abbreviation, but can adjust other parameters as desired like setting splitSize to 0.8 for a holdouts test

```{r}
standardizedModelBuilding <- function(layersVector, cancerName="HNSC", splitSize=0.999, testOnCompleteData=FALSE, graphTitle="", methodNames=c("glmnet"), seed=43, columnsToKeep=c("Age", "gender", "race", "HPV.status", "submitter_id"), stratifying=FALSE, existingExperimentsList=NULL, iterationCount=NULL, applyBiasCorrection=FALSE){
  if (splitSize > 0.98){ testOnCompleteData=TRUE}
  modelBuildingInput <- performSetup(layersVector, cancerName=cancerName, existingExperimentsList=existingExperimentsList)
  
  # Set up title
  if (nchar(graphTitle) == 0){
    string1 <- paste(layersVector, collapse=", ")
    string2 <- paste(methodNames, collapse=", ")
    if (testOnCompleteData){ string3 <- "Complete Set"}
    else{ string3 <- "Holdouts"}
    graphTitle <- paste(string1, "; ", string2, "; ", string3, sep="")
  }
  
  return(list(ModelBuilding=ModelBuilding(modelBuildingInput, cancerName=cancerName, splitSize=splitSize, testOnCompleteData=testOnCompleteData, graphTitle=graphTitle, methodNames=methodNames, seed=seed, columnsToKeep=columnsToKeep, stratifying=stratifying, iterationCount=iterationCount, applyBiasCorrection=applyBiasCorrection), experimentList=modelBuildingInput$experimentsList))
}

```

### Example Code - Run the model

# Note that you must create correct input variables, running this function is not enough yet

```{r message=FALSE}

# Current Test
# modelOutput <- standardizedModelBuilding(c("RPPA", "miRNA"), cancerName="UCEC")
# modelOutput <- standardizedModelBuilding(c("RPPA", "miRNA"), cancerName="UCEC", splitSize=0.8)

# outFile <- "./ModelBuildingOutput/ElasticNetRNAseqCompleteSet.csv"
# write.csv(modelPredictionsPair[[2]], outFile, row.names = FALSE)

# 3 Layers, 3 ML, 5Bayes, Complete Set, using simplified data
# modelBuildingInput <- performSetup(c("miRNA", "RPPA", "RNAseq"), cancerName=cancerName)
# modelPredictionsPair <- ModelBuilding(modelBuildingInput, splitSize=0.999, testOnCompleteData=TRUE, graphTitle="All Layers, Ensemble, Complete Set", methodNames=c("glmnet", "svmLinear", "ranger"), seed=42, columnsToKeep=c("Age", "gender", "race", "HPV.status", "submitter_id"), stratifying=FALSE, iterationCount=5)

# 3 Layers, ElasticNet, Holdouts, using simplified data
# modelBuildingInput <- performSetup(c("miRNA", "RPPA", "RNAseq"), cancerName=cancerName)
# modelPredictionsPair <- ModelBuilding(modelBuildingInput, splitSize=0.8, testOnCompleteData=FALSE, graphTitle="miRNA, RPPA, RNAseq, ElNet, Holdouts", methodNames=c("glmnet"), seed=42, columnsToKeep=c("Age", "gender", "race", "HPV.status", "submitter_id"), stratifying=FALSE)

# # 3 Layers, ElasticNet, complete data set, using simplified data
# modelBuildingInput <- performSetup(c("miRNA", "RPPA", "RNAseq"), cancerName=cancerName)
# modelPredictionsPair <- ModelBuilding(modelBuildingInput, splitSize=0.99, testOnCompleteData=TRUE, graphTitle="miRNA, RPPA, RNASeq, ElNet, Complete Set", methodNames=c("glmnet"), seed=43, pca=FALSE, columnsToKeep=c("Age", "gender", "race", "HPV.status", "submitter_id"), stratifying=FALSE)

# # miRNA, RBF SVM, holdouts
# modelBuildingInput <- performSetup(c("miRNA"), cancerName=cancerName)
# modelPredictionsPair <- ModelBuilding(modelBuildingInput, splitSize=0.8, testOnCompleteData=FALSE, graphTitle="miRNA, RBF SVM, Holdouts", methodNames=c("svmRadial"), seed=42, columnsToKeep=c("Age", "gender", "race", "HPV.status", "submitter_id"), stratifying=FALSE)

# miRNA, ElasticNet, complete set
# modelBuildingInput <- performSetup(c("miRNA"), cancerName=cancerName)
# modelPredictionsPair <- ModelBuilding(modelBuildingInput, splitSize=0.999, testOnCompleteData=TRUE, graphTitle="miRNA, ElasticNet, Complete Set", methodNames=c("glmnet"), seed=43, columnsToKeep=c("Age", "gender", "race", "HPV.status", "submitter_id"), stratifying=FALSE)

# # miRNA, ElasticNet, holdouts
# modelBuildingInput <- performSetup(c("miRNA"), cancerName=cancerName)
# modelPredictionsPair <- ModelBuilding(modelBuildingInput, splitSize=0.8, testOnCompleteData=FALSE, graphTitle="miRNA, ElasticNet, Holdouts", methodNames=c("glmnet"), seed=43, columnsToKeep=c("Age", "gender", "race", "HPV.status", "submitter_id"), stratifying=FALSE)

# # RPPA, ElNet, Complete Set
# modelBuildingInput <- performSetup(c("RPPA"), cancerName=cancerName)
# modelPredictionsPair <- ModelBuilding(modelBuildingInput, splitSize=0.999, testOnCompleteData=TRUE, graphTitle="RPPA, ElasticNet, Complete Set", methodNames=c("glmnet"), seed=43, columnsToKeep=c("Age", "gender", "race", "HPV.status", "submitter_id"), stratifying=FALSE)

# # RPPA, ElNet, holdouts
# modelBuildingInput <- performSetup(c("RPPA"), cancerName=cancerName)
# modelPredictionsPair <- ModelBuilding(modelBuildingInput, splitSize=0.8, testOnCompleteData=FALSE, graphTitle="RPPA, ElasticNet, Holdouts", methodNames=c("glmnet"), seed=43, columnsToKeep=c("Age", "gender", "race", "HPV.status", "submitter_id"), stratifying=FALSE)

# # RNAseq, ElasticNet, complete set, using simplified data
# modelBuildingInput <- performSetup(c("RNAseq"), cancerName=cancerName)
# modelPredictionsPair <- ModelBuilding(modelBuildingInput, splitSize=0.999, testOnCompleteData=FALSE, graphTitle="RNAseq, ElasticNet, Complete Set", methodNames=c("glmnet"), seed=42, columnsToKeep=c("Age", "gender", "race", "HPV.status", "submitter_id"), stratifying=FALSE)

# # RNAseq, ElasticNet, 20% holdouts, using simplified data
# modelBuildingInput <- performSetup(c("RNAseq"), cancerName=cancerName)
# modelPredictionsPair <- ModelBuilding(modelBuildingInput, splitSize=0.8, testOnCompleteData=FALSE, graphTitle="RNAseq, ElasticNet, Holdouts", methodNames=c("glmnet"), seed=42, columnsToKeep=c("Age", "gender", "race", "HPV.status", "submitter_id"), stratifying=FALSE)

# # RNAseq, Random Forest, 20% holdouts, using simplified data
# modelBuildingInput <- performSetup(c("RNAseq"), cancerName=cancerName)
# modelPredictionsPair <- ModelBuilding(modelBuildingInput, splitSize=0.8, testOnCompleteData=FALSE, graphTitle="RNAseq, Random Forest, Holdouts", methodNames=c("ranger"), seed=42, columnsToKeep=c("Age", "gender", "race", "HPV.status", "submitter_id"), stratifying=FALSE)

# # Methylation, Elastic Net, complete data set, using simplified data
# modelBuildingInput <- performSetup(c("Methylation"), cancerName=cancerName)
# modelPredictionsPair <- ModelBuilding(modelBuildingInput, splitSize=0.999, testOnCompleteData=TRUE, seed=43, graphTitle="Methylation, ElasticNet, Complete Set", methodNames=c("glmnet"), columnsToKeep=c("Age", "gender", "race", "HPV.status", "submitter_id"))

## Methylation, Elastic Net, holdouts, using simplified data
# modelBuildingInput <- performSetup(c("Methylation"), cancerName=cancerName)
# modelPredictionsPair <- ModelBuilding(modelBuildingInput, splitSize=0.8, testOnCompleteData=FALSE, seed=43, graphTitle="Methylation, ElasticNet, Holdouts", methodNames=c("glmnet"), columnsToKeep=c("Age", "gender", "race", "HPV.status", "submitter_id"))
```

## Test Bias correction

```{r message=FALSE}
# # RNAseq, ElasticNet, complete set, using simplified data
# modelBuildingInput <- performSetup(c("RNAseq"), cancerName="HNSC")
# modelPredictionsPair <- ModelBuilding(modelBuildingInput, splitSize=0.999, testOnCompleteData=FALSE, graphTitle="RNAseq, ElasticNet, Complete Set", methodNames=c("glmnet"), seed=42, columnsToKeep=c("Age", "gender", "race", "HPV.status", "submitter_id"), stratifying=FALSE, applyBiasCorrection=TRUE)
```

```{r message=FALSE}
# # RPPA, ElNet, Complete Set
# modelBuildingInput <- performSetup(c("RPPA"), cancerName="HNSC")
# modelPredictionsPair <- ModelBuilding(modelBuildingInput, splitSize=0.999, testOnCompleteData=TRUE, graphTitle="RPPA, ElasticNet, Complete Set", methodNames=c("glmnet"), seed=43, columnsToKeep=c("Age", "gender", "race", "HPV.status", "submitter_id"), stratifying=FALSE, applyBiasCorrection=TRUE)
```

```{r message=FALSE}
# # miRNA, ElasticNet, complete set
modelBuildingInput <- performSetup(c("miRNA"), cancerName="HNSC")
modelPredictionsPair <- ModelBuilding(modelBuildingInput, splitSize=0.999, testOnCompleteData=TRUE, graphTitle="miRNA, ElasticNet, Complete Set", methodNames=c("glmnet"), seed=43, columnsToKeep=c("Age", "gender", "race", "HPV.status", "submitter_id"), stratifying=FALSE, applyBiasCorrection=TRUE)
```

```{r message=FALSE}
# # 3 Layers, ElasticNet, complete data set, using simplified data
modelBuildingInput <- performSetup(c("miRNA", "RPPA", "RNAseq"), cancerName="HNSC")
modelPredictionsPair <- ModelBuilding(modelBuildingInput, splitSize=0.99, testOnCompleteData=TRUE, graphTitle="miRNA, RPPA, RNASeq, ElNet, Complete Set", methodNames=c("glmnet"), seed=43, pca=FALSE, columnsToKeep=c("Age", "gender", "race", "HPV.status", "submitter_id"), stratifying=FALSE,  applyBiasCorrection=TRUE)
```
