---
title: "R Notebook"
output: html_notebook
---

```{r message=FALSE, warning=FALSE, include=FALSE}
library(e1071) 
library(randomForest) 
library(SummarizedExperiment)
library(Biobase)
library(ggplot2)
library(caret)
library(survival)
library ("dplyr")
source("PanClockHelperFunctions.R")
source("run_analysis_pipeline.R")
```

### Imputation (needed for methylation)

# This block performs KNN imputation to fill in missing values in the data using five nearest neighbors.

# Note that it runs slowly and may be needed just for methylation, so try to run only as needed and once

# Will skip and return unchanged if impute=TRUE

```{r}
# KNN imputation
# Uses five closest neighbors
imputeMissingValues <- function(experimentsList, assayNames, needsImputation, cancerName, imputationDir="/restricted/projectnb/agedisease/projects/pancancer_aging_clocks/scripts/GitCore/dataset/imputations") {
  print("Checking for existing imputed datasets...")
  imputationDir <- "/restricted/projectnb/agedisease/projects/pancancer_aging_clocks/scripts/imputations_0.05"

  for (i in seq_along(experimentsList)) {
    layerName <- names(experimentsList)[i]
    if (!needsImputation[[layerName]]) next  # Skip if imputation is not needed
    
    experiment <- experimentsList[[layerName]]
    assayName <- assayNames[[layerName]]

    # Define imputation file path: dataset/imputations/{LayerName}/TCGA-{cancerName}_{LayerName}_imputed.rds
    imputationPath <- file.path(imputationDir, layerName, paste0("TCGA-", cancerName, "_", layerName, "_imputed.rds"))
    
    # Check if the imputed file exists
    if (file.exists(imputationPath)) {
      print(paste("Loading existing imputed dataset for", layerName, "(", cancerName, ") from", imputationPath))
      experimentsList[[layerName]] <- readRDS(imputationPath)
    } else {
      print(paste("Imputing missing values for", layerName, "(", cancerName, ") and saving to", imputationPath))
      
      # Perform KNN imputation
      k <- 5
      invisible(capture.output(data_imputed <- impute::impute.knn(as.matrix(SummarizedExperiment::assay(experiment, assayName)), k = k)$data))
      SummarizedExperiment::assay(experiment, paste0(assayName, "-imputed")) <- as.data.frame(data_imputed)

      # Ensure directory exists before saving
      dir.create(dirname(imputationPath), recursive = TRUE, showWarnings = FALSE)
      
      # Save the imputed dataset
      saveRDS(experiment, imputationPath)
      experimentsList[[layerName]] <- experiment
    }
  }
  
  return(experimentsList)
}
```

### Combine Experiments (for doing pan-cancer, which we probably won't)
```{r}
combineExperiments <- function(experimentsList, assayNames){
  match.cols <- Reduce(function(x, y){intersect(x, y@colData@rownames)}, experimentsList, init = colnames(experimentsList[[1]]))
  for (i in 1:(length(experimentsList) - 1)){
    layerName <- names(experimentsList)[i]
    experiment <- experimentsList[[layerName]] 
    experimentsList[[layerName]] <- experiment[,match.cols]
  }
  
  combinedAssays <- assay(experimentsList[[1]], assayNames[[1]])
  for (i in 2:length(experimentsList)){
    combinedAssays <- rbind(combinedAssays, assay(experimentsList[[i]], assayNames[[i]]))
    combinedExperiment <- SummarizedExperiment(assays=combinedAssays, colData = colData(experimentsList[[1]]))
  }
  return(combinedExperiment)
}
# # Remove rows with subjects that don't match
# for (i in 1:(length(experimentsList) - 1)){
#   for (j in (i+1):length(experimentsList)){
#     if (i == j){ next }
#     experiment1 <- experimentsList[[i]]
#     experiment2 <- experimentsList[[j]]
#     experimentsList[[i]] <- experiment1[, sapply(experiment1$submitter_id, checkIfNameMatch, experiment=experiment2)]
#     experimentsList[[j]] <- experiment2[, sapply(experiment2$submitter_id, checkIfNameMatch, experiment=experiment1)]
#   }
# }
```

### Combine the transposed assays of different SummarizedExperiments (for when more than one layer is included)
```{r}
combineAssays <- function(assaysList, combiningNormal=FALSE){
  print("Combining assays...")
  combinedAssays <- assaysList[[1]]
  if (length(assaysList) == 1){ return(combinedAssays) }
  for (i in 2:length(assaysList)){
    if (combiningNormal){
      print("Taking differences with normal...")
      normalAssay <- assaysList[[i]]
      combinedAssays <- combinedAssays %>% filter(submitter_id %in% normalAssay$submitter_id)
      normalAssay <- normalAssay %>% filter(submitter_id %in% combinedAssays$submitter_id)
      combinedAssays <- combinedAssays[ , intersect(names(combinedAssays), names(normalAssay))]
      normalAssay <- normalAssay[ , intersect(names(normalAssay), names(combinedAssays))]
      numericColumns <- sapply(combinedAssays, is.numeric)
      combinedAssays[numericColumns] <- combinedAssays[numericColumns] - normalAssay[numericColumns]
    }
    else{
      combinedAssays <- merge(combinedAssays, assaysList[[i]], by="submitter_id", all=FALSE)
    }
    combinedAssays$Row.names <- NULL
  }
  return(combinedAssays)
}
```

# Remove subjects missing covariates from experiments list and collect covariates
```{r}
# Only keep certain metadata (determined in 03_Covariates.html)
filterForCovariates <- function(experimentsList, featuresToEncode=c("gender"), columns_to_keep=c("Age", "vital_status", "Survival_Time", "submitter_id")){
  print("Filtering for covariates...")
  covariatesList <- list()
  for (i in 1:length(experimentsList)){
    layerName <- names(experimentsList)[i]
    experiment <- experimentsList[[layerName]]
    covariates <- SummarizedExperiment::colData(experiment)[, c(featuresToEncode, columns_to_keep), drop = FALSE]

    # Check for and remove rows with missing values
    covariates <- covariates[complete.cases(as.data.frame(covariates)), ]
    
    # Filter layer data to match samples
    layer_processed <- experiment[, rownames(covariates)]
    experimentsList[[layerName]] <- layer_processed
    covariatesList[[layerName]] <- covariates
  }
  return(list(experiments=experimentsList, covariates=covariatesList))
}
```

### Gets assay data using pre-filtered data if available as well as imputed vals.
```{r}
getCleanedAssayData <- function(experimentsList, assayNames, useImputed, simplifiedDataPairs=list(), significanceCutoff=0.025){
  print("Getting cleaned assay data...")

  # Only keep rows that are significantly age associated based on prior calculations
  assayList <- list()
  for (i in 1:length(experimentsList)){    
    layerName <- names(experimentsList)[i]
    experiment <- experimentsList[[layerName]]
    if (length(simplifiedDataPairs) > 0 && length(simplifiedDataPairs[[layerName]]) > 1){
      simplifiedData <- simplifiedDataPairs[[layerName]]$data
      significantFeatures <- simplifiedData[simplifiedData[[simplifiedDataPairs[[layerName]]$metric]] <= significanceCutoff, ]
      if (assayNames[[layerName]] == "M-values"){ 
        experiment <- experiment[rownames(significantFeatures), ] }
      else{ experiment <- experiment[rowRanges(experiment)$gene_name %in% rownames(significantFeatures), ] }
    }
  
    # Return the assay data for the experiment, using the imputed values if available
    if (useImputed[[layerName]]){ assayType <- paste(assayNames[[layerName]],"-imputed",sep="")}
    else { assayType <- assayNames[[layerName]]}
    assayData <- getTransposedFrame(assay(experiment, assayType))
    assayData <- cbind(assayData, submitter_id=row.names(assayData))
    assayList[[layerName]] <- assayData
  }
  return(assayList)
}
```

### Do any additional feature filtering here; necessity may depend on ML method
```{r}
filterFeatures <- function(assayList, experimentsList, relativeVariance=FALSE, highCorrelationToAge=FALSE, highCorrelationToEachOther=FALSE, cutoffThreshold=0.99){
  print("Additional feature filtering...")
  for (i in 1:length(assayList)){
    layerName <- names(assayList)[i]
    assayData <- assayList[[layerName]]
    
    if (relativeVariance){
      assayData <- assayData[, apply(assayData, 2, isRelativeVarianceHigh)]
    }
    if (highCorrelationToAge){
      assayData <- assayData[, cor(experimentsList[[layerName]]$Age, assayData) > 0.001]
    }
    # Remove features highly correlated with each other
    if (highCorrelationToEachOther){
      dataLeft <- TRUE
      i <- 1
      while(dataLeft){
        after_i <- seq(i+1, ncol(assayData))
        correlations <- cor(assayData[,after_i], assayData[,i])
        if(any(abs(correlations) > cutoffThreshold)){
          assayData <- assayData[, -after_i[which(abs(correlations) > cutoffThreshold)]]
        }
        i <- i+1
        if(ncol(assayData) <= i){ dataLeft <- FALSE }
      }
    }
    
    assayList[[layerName]] <- assayData
  }
  return(assayList)
}
```

### Set up one-hot encoding of covariates
```{r}
encode_covariates <- function(covariates, assayData, featuresToEncode=c(), columns_to_keep=c("Age", "vital_status", "Survival_Time", "submitter_id"), encodingHPV=FALSE, encodingGender=FALSE, encodingRace=FALSE){
  print("Encoding covariates...")

  # One-hot encoding for gender, race, and potentially HPV.status
  # if (nlevels(as.factor(covariates$race)) > 1){ encodingRace <- TRUE }
  if (nlevels(as.factor(covariates$gender)) > 1){ encodingGender <- TRUE }

  # Build up formula depending on available covariates
  if (length(featuresToEncode) == 0){ encoded_covariates <- covariates }
  else{
    firstCovariateFound <- FALSE
    for (feature in featuresToEncode){
      if (nlevels(as.factor(covariates[[feature]])) > 1){ # Check if covariate has more than 1 label

        # Initialize formula
        if (!firstCovariateFound){
          covariateFormula = paste("~", feature)
          firstCovariateFound <- TRUE
        }
        
        # Add more covariates
        else{ covariateFormula = paste(covariateFormula, "+", feature) }
      }
    }
    if (!firstCovariateFound){ encoded_covariates <- as.data.frame(covariates[, colnames(covariates) %in% columns_to_keep]) }
    else{
      # Set up one hot encoded covariates based on derived formula
      dummies <- caret::dummyVars(covariateFormula, data = covariates)
      covariateFormula <- paste(covariateFormula, "- 1")
      encoded_covariates <- data.frame(model.matrix(as.formula(covariateFormula), data = covariates))
      for (col in columns_to_keep){ encoded_covariates[[col]] <- covariates[[col]] }
    }
  }
  
  # Combine encoded covariates with assay data
  rownames(encoded_covariates) <- encoded_covariates$submitter_id
  assayDataWithCovariates <- merge(encoded_covariates, assayData, by="submitter_id", all=FALSE)
  rownames(assayDataWithCovariates) <- assayDataWithCovariates$submitter_id
  assayDataWithCovariates$Row.names <- NULL
  
  return(assayDataWithCovariates)
}
```

# Adjust tumor assay by sd and mean of each feature in its non-tumor assay. Currently only set up for one layer at a time
```{r}
scaleAssayByNormal <- function(experimentsList, assayNames, needsImputation, cancerName){
  
  print("Scaling assay...")
  
  # Initialize parameters and nor
  tumorExperiment <- experimentsList[[1]]
  assayName <- assayNames[[1]]
  layerName <- names(experimentsList)[1]
  if (needsImputation[[1]]){ tumorAssayName <- paste(assayName, "-imputed", sep="") }

  # Set normal experiment
  normalLayerName <- paste(layerName, "_Normal", sep="")
  normalExperiment <- getExperimentByLayerAndCancer(normalLayerName, cancerName)[[normalLayerName]]
  
  # Set experiments to have same features
  tumorExperiment <- reduceExperimentByFeatures(tumorExperiment, normalExperiment, tumorAssayName, assayName2=assayName)
  normalExperiment <- reduceExperimentByFeatures(normalExperiment, tumorExperiment, assayName, assayName2=tumorAssayName)
  
  # Scale tumor assay
  normAssay <- getTransposedFrame(assay(normalExperiment, assayName))
  tumorAssay <- getTransposedFrame(assay(tumorExperiment, tumorAssayName))
  means <- colMeans(normAssay)
  sds <- sapply(normAssay, sd)
  tumorAssay <- scale(tumorAssay, center = means, scale = sds)

  # Set assay in tumor experiment
  SummarizedExperiment::assay(tumorExperiment, tumorAssayName) <- getTransposedFrame(tumorAssay)
  experimentsList[[1]] <- tumorExperiment
  return(experimentsList)
}
```