---
title: "R Notebook"
output: html_notebook
---

```{r message=FALSE, warning=FALSE, include=FALSE}
library(e1071) 
library(randomForest) 
library(SummarizedExperiment)
library(Biobase)
library(ggplot2)
library(caret)
library ("dplyr")
source("PanClockHelperFunctions.R")
source("run_analysis_pipeline.R")
```

### Imputation (needed for methylation)

# This block performs KNN imputation to fill in missing values in the data using five nearest neighbors.

# Note that it runs slowly and may be needed just for methylation, so try to run only as needed and once

# Will skip and return unchanged if impute=TRUE

```{r}
#KNN imputation
#Use five closest neighbors
imputeMissingValues <- function(experimentsList, assayNames, needsImputation){
  print("Imputing...")
  for (i in 1:length(experimentsList)){
    layerName <- names(experimentsList)[i]
    print(layerName)
    if (needsImputation[[layerName]]){
      experiment <- experimentsList[[layerName]]
      assayName <- assayNames[[layerName]]
      k = 5
      invisible(capture.output(data_imputed <- impute::impute.knn(as.matrix(SummarizedExperiment::assay(experiment,assayName)), k = k )$data))
      SummarizedExperiment::assay(experiment,paste(assayName, "-imputed", sep="")) <- as.data.frame(data_imputed)
      experimentsList[[layerName]] <- experiment
    }
  }
  return(experimentsList)
}
```

### Combine Experiments (for doing pan-cancer, which we probably won't)
```{r}
combineExperiments <- function(experimentsList, assayNames){
  match.cols <- Reduce(function(x, y){intersect(x, y@colData@rownames)}, experimentsList, init = colnames(experimentsList[[1]]))
  for (i in 1:(length(experimentsList) - 1)){
    layerName <- names(experimentsList)[i]
    experiment <- experimentsList[[layerName]] 
    experimentsList[[layerName]] <- experiment[,match.cols]
  }
  
  combinedAssays <- assay(experimentsList[[1]], assayNames[[1]])
  for (i in 2:length(experimentsList)){
    combinedAssays <- rbind(combinedAssays, assay(experimentsList[[i]], assayNames[[i]]))
    combinedExperiment <- SummarizedExperiment(assays=combinedAssays, colData = colData(experimentsList[[1]]))
  }
  return(combinedExperiment)
}
# # Remove rows with subjects that don't match
# for (i in 1:(length(experimentsList) - 1)){
#   for (j in (i+1):length(experimentsList)){
#     if (i == j){ next }
#     experiment1 <- experimentsList[[i]]
#     experiment2 <- experimentsList[[j]]
#     experimentsList[[i]] <- experiment1[, sapply(experiment1$submitter_id, checkIfNameMatch, experiment=experiment2)]
#     experimentsList[[j]] <- experiment2[, sapply(experiment2$submitter_id, checkIfNameMatch, experiment=experiment1)]
#   }
# }
```

### Combine the transposed assays of different SummarizedExperiments (for when more than one layer is included)
```{r}
combineAssays <- function(assaysList, combiningNormal=FALSE){
  print("Combining assays...")
  combinedAssays <- assaysList[[1]]
  if (length(assaysList) == 1){ return(combinedAssays) }
  for (i in 2:length(assaysList)){
    if (combiningNormal){
      print("Taking differences with normal...")
      normalAssay <- assaysList[[i]]
      combinedAssays <- combinedAssays %>% filter(submitter_id %in% normalAssay$submitter_id)
      normalAssay <- normalAssay %>% filter(submitter_id %in% combinedAssays$submitter_id)
      combinedAssays <- combinedAssays[ , intersect(names(combinedAssays), names(normalAssay))]
      normalAssay <- normalAssay[ , intersect(names(normalAssay), names(combinedAssays))]
      numericColumns <- sapply(combinedAssays, is.numeric)
      combinedAssays[numericColumns] <- combinedAssays[numericColumns] - normalAssay[numericColumns]
    }
    else{
      combinedAssays <- merge(combinedAssays, assaysList[[i]], by="submitter_id", all=FALSE)
    }
    combinedAssays$Row.names <- NULL
  }
  return(combinedAssays)
}
```

# Remove subjects missing covariates from experiments list and collect covariates
```{r}
# Only keep certain metadata (determined in 03_Covariates.html)
filterForCovariates <- function(experimentsList, columns_to_keep=c("Age", "gender", "race", "HPV.status", "submitter_id")){
  print("Filtering for covariates...")
  covariatesList <- list()
  for (i in 1:length(experimentsList)){
    layerName <- names(experimentsList)[i]
    experiment <- experimentsList[[layerName]]
    covariates <- SummarizedExperiment::colData(experiment)[, columns_to_keep, drop = FALSE]
    # Check for and remove rows with missing values
    covariates <- covariates[complete.cases(covariates), ]
    # Filter layer data to match samples
    layer_processed <- experiment[, rownames(covariates)]
    experimentsList[[layerName]] <- layer_processed
    covariatesList[[layerName]] <- covariates
  }
  return(list(experiments=experimentsList, covariates=covariatesList))
}
```

### Gets assay data using pre-filtered data if available as well as imputed vals.
```{r}
getCleanedAssayData <- function(experimentsList, assayNames, useImputed, simplifiedDataPairs=list()){
  print("Getting cleaned assay data...")

  # Only keep rows that are significantly age associated based on prior calculations
  assayList <- list()
  for (i in 1:length(experimentsList)){    
    layerName <- names(experimentsList)[i]
    experiment <- experimentsList[[layerName]]
    if (length(simplifiedDataPairs) > 0 && length(simplifiedDataPairs[[layerName]]) > 1){
      simplifiedData <- simplifiedDataPairs[[layerName]]$data
      significantFeatures <- simplifiedData[simplifiedData[[simplifiedDataPairs[[layerName]]$metric]] <= 0.025, ]
      if (assayNames[[layerName]] == "M-values"){ 
        experiment <- experiment[rownames(significantFeatures), ] }
      else{ experiment <- experiment[rowRanges(experiment)$gene_name %in% rownames(significantFeatures), ] }
    }
  
    # Return the assay data for the experiment, using the imputed values if available
    if (useImputed[[layerName]]){ assayType <- paste(assayNames[[layerName]],"-imputed",sep="")}
    else { assayType <- assayNames[[layerName]]}
    assayData <- as.data.frame(t(assay(experiment, assayType)))
    assayData <- cbind(assayData, submitter_id=row.names(assayData))
    assayList[[layerName]] <- assayData
  }
  return(assayList)
}
```

### Do any additional feature filtering here; necessity may depend on ML method
```{r}
filterFeatures <- function(assayList, experimentsList, relativeVariance=FALSE, highCorrelationToAge=FALSE, highCorrelationToEachOther=FALSE, cutoffThreshold=0.99){
  print("Additional feature filtering...")
  for (i in 1:length(assayList)){
    layerName <- names(assayList)[i]
    assayData <- assayList[[layerName]]
    
    if (relativeVariance){
      assayData <- assayData[, apply(assayData, 2, isRelativeVarianceHigh)]
    }
    if (highCorrelationToAge){
      assayData <- assayData[, cor(experimentsList[[layerName]]$Age, assayData) > 0.001]
    }
    # Remove features highly correlated with each other
    if (highCorrelationToEachOther){
      dataLeft <- TRUE
      i <- 1
      while(dataLeft){
        after_i <- seq(i+1, ncol(assayData))
        correlations <- cor(assayData[,after_i], assayData[,i])
        if(any(abs(correlations) > cutoffThreshold)){
          assayData <- assayData[, -after_i[which(abs(correlations) > cutoffThreshold)]]
        }
        i <- i+1
        if(ncol(assayData) <= i){ dataLeft <- FALSE }
      }
    }
    
    assayList[[layerName]] <- assayData
  }
  return(assayList)
}
```

### Set up one-hot encoding of covariates. Currently designed for one set of assay data

```{r}
encode_covariates <- function(covariates, assayData, featuresToEncode=NULL, encodingHPV=FALSE, encodingGender=FALSE, encodingRace=FALSE){
  print("Encoding covariates...")

    # One-hot encoding for gender, race, and potentially HPV.status
  if (nlevels(as.factor(covariates$race)) > 1){ encodingRace <- TRUE }
  if (nlevels(as.factor(covariates$gender)) > 1){ encodingGender <- TRUE }
  
  # Return early if no suitable covariates
  if (!encodingRace && !encodingGender && !encodingHPV){ return(assayData) }
  
  # Build up formula depending on available covariates
  covariateFormula = "~"
  if (encodingGender){ covariateFormula <- paste(covariateFormula, "gender") }
  if (encodingRace){ 
    if (encodingGender){ covariateFormula <- paste(covariateFormula, "+") }
    covariateFormula <- paste(covariateFormula, "race") 
  }
  if (encodingHPV){
    if (encodingGender || encodingRace){ covariateFormula <- paste(covariateFormula, "+") }
    covariateFormula <- paste(covariateFormula, "HPV.status") 
  }
  
  dummies <- caret::dummyVars(covariateFormula, data = covariates)
  covariateFormula <- paste(covariateFormula, "- 1")
  encoded_covariates <- data.frame(model.matrix(as.formula(covariateFormula), data = covariates))
  encoded_covariates$Age <- covariates$Age
  encoded_covariates$submitter_id <- covariates$submitter_id
  rownames(encoded_covariates) <- encoded_covariates$submitter_id

  assayDataWithCovariates <- merge(encoded_covariates, assayData, by="submitter_id", all=FALSE)
  rownames(assayDataWithCovariates) <- assayDataWithCovariates$submitter_id
  assayDataWithCovariates$Row.names <- NULL
  return(assayDataWithCovariates)
}
```
